---
output: word_document
---

#Title: "Storm Data Analysis"


##Synopsis:

This project involves exploring the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database. The analisys allows to answer which types of events are most harmful with respect to population health and which types of events have the greatest economic consequences

##Data processing

The data for this assignment come in the form of a comma-separated-value file compressed via the bzip2 algorithm to reduce its size. The events in the database start in the year 1950 and end in November 2011

```{r, cache=TRUE, warning=FALSE}
library(R.utils)
bunzip2("repdata_data_StormData.csv.bz2", "StormData.csv", remove = FALSE, skip = TRUE)
storm <- read.csv("StormData.csv")
storm$BGN_DATE<- as.Date(storm$BGN_DATE, "%m/%d/%Y")
storm$YEAR<- as.numeric(format(storm$BGN_DATE, "%Y"))
```

Selecting the correct characteristics for data analysis
From the NOAA data, the most important data for this analysis are YEAR 
EVTYPE, FATALITIES, INJURIES, PROPDMG, and CROPDMG. 

```r
storm <- storm[,c("YEAR","EVTYPE","FATALITIES","INJURIES","PROPDMG","CROPDMG")]
dim(storm)
```

Fixing the most common mistakes in the Severe Weathter Event Types

```r
storm$EVTYPE <- toupper(storm$EVTYPE)
storm$EVTYPE = gsub("TSTM","THUNDERSTORM",storm$EVTYPE)
storm$EVTYPE = gsub("URBAN/SML STREAM FLD","THUNDERSTORM WIND",storm$EVTYPE)
storm$EVTYPE = gsub("TSTM WIND","THUNDERSTORM WIND",storm$EVTYPE)
storm$EVTYPE = gsub("THUNDERSTORM/HAIL","THUNDERSTORM WIND",storm$EVTYPE)
storm$EVTYPE = gsub("RIP CURRENTS","RIP CURRENT",storm$EVTYPE)
storm$EVTYPE = gsub("TYPHOON","HURRICANE",storm$EVTYPE)
storm$EVTYPE = gsub("TEMPERATURE RECORD","HEAT",storm$EVTYPE)
storm$EVTYPE = gsub("SMALL HAIL","HAIL",storm$EVTYPE)
storm$EVTYPE = gsub("THUNDERSTORMS","THUNDERSTORM",storm$EVTYPE)
storm$EVTYPE = gsub("HURRICANE/TYPHOON","HURRICANE",storm$EVTYPE)
storm$EVTYPE = gsub("HURRICANE/HURRICANE","HURRICANE",storm$EVTYPE)
storm$EVTYPE = gsub("STORM SURGE/TIDE","STORM SURGE",storm$EVTYPE)
storm$EVTYPE = gsub("VOLCANIC ASHFALL","VOLCANIC ASH",storm$EVTYPE)
storm$EVTYPE = gsub("   HIGH SURF ADVISORY","HIGH SURF",storm$EVTYPE)
storm$EVTYPE = gsub("HEAVY SURF/HIGH SURF","HIGH SURF",storm$EVTYPE)
storm$EVTYPE = gsub("$HAIL","HAIL", storm$EVTYPE)
storm$EVTYPE = gsub("(TORNADO|^TORNADO)","TORNADO", storm$EVTYPE)
storm$EVTYPE = gsub("(FLOOD|^FLOOD)","FLOOD", storm$EVTYPE)
storm$EVTYPE = gsub("(WARM|^WARM)","HEAT", storm$EVTYPE)
storm$EVTYPE = gsub("COLD","EXTREME COLD/WIND CHILL", storm$EVTYPE)
storm$EVTYPE = gsub("(SLEET|^SLEET)","SLEET", storm$EVTYPE)
storm$EVTYPE = gsub("(ICE|^ICE)","FROST/FREEZE", storm$EVTYPE)
storm$EVTYPE = gsub("STRONG WINDS","STRONG WIND", storm$EVTYPE)
```


##Results 

### Question 1: Across the United States, which types of events (as indicated in the EVTYPE variable) are most harmful with respect to population health? 

Analyzing the deaths caused by Severe Weather Event

```{r, echo=TRUE}
deathsF <- aggregate(FATALITIES ~ EVTYPE, data = storm, FUN = sum)
deathsF <- deathsF[order(-deathsF$FATALITIES), ]
head(deathsF,10)
```

As we can see, TORNADO has the greatest number of fatalities


Analyzing the injuries caused by Severe Weather Event

```{r, echo=TRUE}
deathsI <- aggregate(INJURIES ~ EVTYPE, data = storm, FUN = sum)
deathsI <- deathsI[order(-deathsI$INJURIES), ]
head(deathsI,10)
```

Then, TORNADO has the greatest number in causing deaths and injuries 

Now, it's created a bar using ggplot to show the most harmful Severe Weather Events Type for Population Health

```{r, echo=TRUE, warning=FALSE}
fatalities_injuries <- aggregate(cbind(FATALITIES + INJURIES) ~ EVTYPE, data = storm, FUN = sum)
colnames(fatalities_injuries)<- c("Event", "Total")
fatalities_injuries <- fatalities_injuries[order(-fatalities_injuries$Total),]
fatalities_injuries<- head(fatalities_injuries, 10)

library(ggplot2)
g <- ggplot(fatalities_injuries,aes(Event,Total, fill = fatalities_injuries$Event)) +
    geom_bar(stat="identity",width=0.75) +    
    labs(x="Severe weather Event Type", y="Total number of Events") + 
    guides(fill = FALSE)+
    labs(title="Top 10 Most Harmful Events to Population Health")+
    theme(axis.text.x = element_text(angle=45, vjust=0.5, size=10))

print(g)
```

From the bar plot above, we can see that TORNADO, THUNDERSTORM WIND and FLASH FLOOD Were the top three most harmful events for Population Health.

### Question 2: Across the United States, which types of events have the greatest economic consequences?

Analyzing the crop damage caused by Severe Weather Event

```{r, echo=TRUE}
crop_dam <- aggregate(CROPDMG ~ EVTYPE, data = storm, FUN = sum)
crop_dam <- crop_dam[order(-crop_dam$CROPDMG), ]
head(crop_dam,10)
```

As we can see HAIL has the greatest number in crop damage

Analyzing the property damage caused by Severe Weather Event

```{r, echo=TRUE}
prop_dam <- aggregate(PROPDMG ~ EVTYPE, data = storm, FUN = sum)
prop_dam <- prop_dam[order(-prop_dam$PROPDMG), ]
head(prop_dam,10)
```

TORNADO has the greatest number in property damage

Now, it's created a bar using ggplot to show which types of events have the greatest economic consequences

```{r, echo=TRUE}
crop_prop_damage <- aggregate(cbind(CROPDMG + PROPDMG) ~ EVTYPE, data = storm, FUN = sum)
colnames(crop_prop_damage)<- c("Event", "Total")
crop_prop_damage$Total <- crop_prop_damage$Total/1000000000
crop_prop_damage <- crop_prop_damage[order(-crop_prop_damage$Total),]
crop_prop_damage<- head(crop_prop_damage, 10)

g1 <- ggplot(crop_prop_damage,aes(Event,Total, fill = crop_prop_damage$Event)) +
    geom_bar(stat="identity",width=0.75) +    
    labs(x="Severe weather Event Type", y="Total amount of damage(in billions of dollars)") + 
    guides(fill = FALSE)+
    labs(title="Top 10 events with the greatest economic consequences")+
    theme(axis.text.x = element_text(angle=45, vjust=0.5, size=10))

print(g1)
```

From the bar plot above, we can see that TORNADO, THUNDERSTORM WIND and FLASH FLOOD Were the top three most harmful events for Population Health.













